Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-02-08 paper/letter.tex
nueva estructura de parrafos

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Lyman-alpha emission reveals an unusual fastly rotating
  compact dwarf galaxy}  
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$,
\\
\\
\scriptsize
{$^1$ Universidad de los Andes
\\
$^2$ Arizona
\normalsize
\\
\\
\textbf{
Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected pristine
conditions of the first galaxies.  
Until these early galaxy generations are observationally
detected, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.  
Here we report on one of such CDGs, \tol, which presents
features in its Lyman-$\alpha$ emission thad had evaded theoretical
interpretation so far. 
We show that these special features, a symmetric triple peaked
emission line, are naturally explained by gas rotation.  
We find that the Lyman-$\alpha$ emission region in \tol\ should
have a rotational velocity of $V_{r}=300$ km s$^{-1}$ and a neutral
Hydrogen column density of $\log N_{HI}=20.5$ atoms cm$^{-2}$.  
Considering other observational information we find that
the diameter for that region should be in the range of
$110$ pc $<D<$ $340$ pc and its total dynamical mass 
should be between $2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times
10^{9}$M$_{\odot}$. 
This dynamical mass is at least $16\pm 9$ times
larger than the neutral mass hydrogen.
We argue that a possibility to explain the excess in
dynamical mass is the presence of a super-massive black hole. }

@@

1. General paragraph about the Lyman alpha line.

2. General paragraph about modelling the Lyman alpha
line. Outflows.

3. Rotation and the expected features.
It has been shown that rotation also imprints an effect on the
Lyman-alpha morphology. 
The most important consequence of rotation is that spherical 
symmetry is broken.
The line morphology now depends on the viewing angle respect to the
rotation axis.  
For a line of sight perpendicular to the rotation axis the intensity and
the line center and the line width increase with rotational velocity. 
When the rotational velocity is close to the half-line width of the
static line, the line becomes single peaked as it is observed in
\tol, a unique feature that other theoretical models find
impossible to reproduce.

%\section{TOL1214-277}

4. The charachteristics of the dwarf galaxy of interest.

@@

The Ly$\alpha$ emission line was first observed in the TOL1214-277 
galaxy by \cite{Thuan97}. It has two main important features which 
make this a very uncommun LAE. First it shown a symmetric profile 
which is , Second the Ly$\alpha$ line is not shifted with respect to 
the H$\beta$ line.  Blue compact Dwarf Galaxy

	%\section{Results}

5. The results of the fit.

@@

Observed line + fit.

%6. Comparison of these results against other observational data.
Assuming spherical symmetry and a homogeneous gas distribution we
estimate the total neutral hydrogen mass to be on the order of
$M\approx m_H\tau^3\sigma^{-3}n^{-2}$, where $m_H$ is the mass a
Hydrogen's atom, $\tau$  the optical depth, $\sigma$ is the cross
section at the line's center and $n$ is the number density of neutral
Hydrogen atoms. 
For this system we estimate that for average values of $n=1\times
10^3$ the total hydrogen mass is $M\times 10^{14}$M$_{\odot}$. 
However, blind HI
surveys have put an upper limit in the neutral hydrogen mass of ???

7. Implications for outflow+rotation in existing samples.

%8. Implications for very high-z dwarf galaxies.

@@

\newpage
\begin{table}
\begin{center}

\begin{tabular}{lc}
$\alpha$(2000)$^{a}$ & 12h17min17.1s\\
$\delta$(2000)$^{b}$ & -28d02m32s\\
$l$, $b$ (deg) & 294, 34\\
$m_V$ & 17.5\\
$v$(km s$^{-1}$) & 7795
\end{tabular}
\end{center}

\caption{Observational characteristics of TOL1214-277 \cite{Thuan97}\\
$^{a}$ Units of right ascension are hours, minutes and
  seconds.\\ $^{b}$ Units of declination are degrees, arcminutes and arcseconds.}
\end{table}

@@

>>>>

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Lyman-alpha emission reveals an unusual fastly rotating
  compact dwarf galaxy}  
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$,
\\
\\
\scriptsize
{$^1$ Universidad de los Andes
\\
$^2$ Arizona
\normalsize
\\
\\
\textbf{
Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected pristine
conditions of the first galaxies.  
Until these early galaxy generations are observationally
detected, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.  
Here we report on one of such CDGs, \tol, which presents
features in its Lyman-$\alpha$ emission thad had evaded theoretical
interpretation so far. 
We show that these special features, a symmetric triple peaked
emission line, are naturally explained by gas rotation.  
We find that the Lyman-$\alpha$ emission region in \tol\ should
have a rotational velocity of $V_{r}=300$ km s$^{-1}$ and a neutral
Hydrogen column density of $\log N_{HI} / \mathrm{atoms\ cm}^{-2} =
  20.5$.   
Using archival observational information we find that
the diameter for that region should be in the range of
$110$ pc $<D<$ $340$ pc and its total dynamical mass 
should be between $2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times
10^{9}$M$_{\odot}$. 
This dynamical mass is at least $16\pm 9$ times
larger than the neutral mass hydrogen.
We argue that a possibility to explain the excess in
dynamical mass is the presence of a super-massive black hole. }

@@

1. [General paragraph about the first galaxies and CDGs]

2. [General paragrph about the expected Lyman-alpha emission]. Almost
fifty years ago \cite{PartridgePeebles} it was realized that young
galaxies should have a strong Lyman-$\alpha$ line emission. 
This theoretical prediciotn was only confirmed XX year later. 
Currently, Lyman Alpha Emitting galaxies are commonly measure in
surveys.

3. [The link betwween high-z galaxies and lyman alpha, LARS]

4. [Observations require interpretation, common features and
  models]. The Lyman-$\alpha$ line is resonant. A yman-$\alpha$ photon
will follow a difussion-like process before escaping the galaxy or
being absorbed by dust. This makes it hard to interpret Lyman-alpha
line observations as the reulting profile is sensitive to the
dynamical, chemical and thermal conditions in the interstellar
medium. State-of-the-art models

5. [Rotation and the expected features.]
It has been shown that rotation also imprints an effect on the
Lyman-alpha morphology. 
The most important consequence of rotation is that spherical 
symmetry is broken.
The line morphology now depends on the viewing angle respect to the
rotation axis.  
For a line of sight perpendicular to the rotation axis the intensity
and the line center and the line width increase with rotational
velocity.  
When the rotational velocity is close to the half-line width of the
static line, the line becomes single peaked as it is observed in
\tol, a unique feature that other theoretical models find
impossible to reproduce.

6. [The charachteristics of the dwarf galaxy of interest as a puzzling
system]

@@

The Ly$\alpha$ emission line was first observed in the \tol
galaxy by \cite{Thuan97}. It has two main important features which 
make this a very uncommun LAE. First it shown a symmetric profile 
which is , Second the Ly$\alpha$ line is not shifted with respect to 
the H$\beta$ line.  Blue compact Dwarf Galaxy

7. [The results of the fit + Physical picture using other results]

@@

Figura Observed line + fit.

%6. Comparison of these results against other observational 
Assuming spherical symmetry and a homogeneous gas distribution we
estimate the total neutral hydrogen mass to be on the order of
$M\approx m_H\tau^3\sigma^{-3}n^{-2}$, where $m_H$ is the mass a
Hydrogen's atom, $\tau$  the optical depth, $\sigma$ is the cross
section at the line's center and $n$ is the number density of neutral
Hydrogen atoms. 
For this system we estimate that for average values of $n=1\times
10^3$ the total hydrogen mass is $M\times 10^{14}$M$_{\odot}$. 
However, blind HI
surveys have put an upper limit in the neutral hydrogen mass of ???

8. [Possible interpretation of this results]

9. [Implications for outflow+rotation in existing samples.]

10. [Future outlook]

@@

\newpage

\section*{\tol\ characteristics}

\begin{table}
\begin{center}
\begin{tabular}{lc}
$\alpha$(2000)$^{a}$ & 12h17min17.1s\\
$\delta$(2000)$^{b}$ & -28d02m32s\\
$l$, $b$ (deg) & 294, 34\\
$m_V$ & 17.5\\
$v$(km s$^{-1}$) & 7795
\end{tabular}
\end{center}

\caption{Observational characteristics of TOL1214-277
  \cite{Thuan97}\\} 
\end{table}

@@

\section*{MCMC constraints}

\section*{Physical Interpretation}
